Network Configuration Synthesis with Abstract Topologies

### 2016-11-06 paper/abstraction/appendix.tex
start adding substitution theorem to the appendix

\onecolumn
\section{Preliminaries}

@@

@@

>>>>

\onecolumn
\section{Definitions}

@@

\subsection{Abstraction}

A network topology is a graph $G$ = ($V, E$), where $V$ is a set of vertices and $E \colon V \times V$ a set of directed edges. A role-based abstraction is a graph homomorphism from $G$ to an abstract graph $G^A$ = ($V^A$,$E^A$). A graph homorphism $f : G \rightarrow G^A$ is a function that maps each node in the concrete graph to a node in the abstract graph such that, whenever $(u,v) \in E$, then $(f(u),f(v)) \in E^A$. We write $f^{-1}(v)$ to denote the set of concrete nodes mapping to the abstract node $v$. Given a path as a sequence of locations $p = l_1, \dots, l_k$ we define $f(p) = f(l_1), \dots, f(l_k)$.

@@

\subsection{Automata}

\newcommand{\Lang}{\ensuremath{\mathcal{L}}}

\begin{lem}
  For any map $m$ and regular expression $r$ over abstract locations, path $p \in \Lang(con(r,m)) \iff f(p) \in \Lang(r)$.
\end{lem}
\begin{proof}
By induction on the structure of $r$

\emph{Case $\emptyset$:}
  \[ \begin{array}{lll}
    p \in \Lang(con(\emptyset)) &\iff& f(p) \in \Lang(\emptyset) \\
    p \in \emptyset &\iff& f(p) \in \emptyset \\
  \end{array} \]

\emph{Case $l$:}
  \[ \begin{array}{lll}
    p \in \Lang(con(l)) &\iff& f(p) \in \Lang(l) \\
    p \in \Lang(\Sigma ~ f^{-1}(l)) &\iff& f(p) \in \Lang(l) \\
    p \in \bigcup f^{-1}(l) &\iff& f(p) \in \Lang(l) \\
    p \in f^{-1}(l) &\iff& f(p) \in \{ l \} \\
    p \in f^{-1}(l) &\iff& f(p) = l \\
    \text{by homomorphism} & & \\
  \end{array} \]

\emph{Case $r_1 \cup r_2$:}
  \[ \begin{array}{lll}
    p \in \Lang(con(r_1 \cup r_1)) &\iff& f(p) \in \Lang(r_1 \cup r_2) \\
    p \in \Lang(con(r_1) \cup con(r_1)) &\iff& f(p) \in \Lang(r_1 \cup r_2) \\
    p \in \Lang(con(r_1)) \cup \Lang(con(r_1)) &\iff& f(p) \in \Lang(r_1 \cup r_2) \\
    p \in \Lang(con(r_1)) \vee p \in \Lang(con(r_1)) 
       &\iff& 
       f(p) \in \Lang(r_1) \vee f(p) \in \Lang(r_2) \\
    \text{by cases + IH} & & \\
  \end{array} \]

\emph{Case $r_1 \cap r_2$:} 
  \[ \begin{array}{c}
    \text{symmetric to } \cup \text{ case}
  \end{array} \]

\emph{Case $!r$:}
  \[ \begin{array}{lll}
    p \in \Lang(con(!r)) &\iff& f(p) \in \Lang(!r) \\
    p \in \Lang(!con(r)) &\iff& f(p) \in \Lang(!r) \\
    p \in \Sigma^* - \Lang(con(r)) &\iff& f(p) \in \Sigma^* - \Lang(r) \\
    \text{by IH} & & \\
  \end{array} \]

\emph{Case $r^*$:}
  \[ \begin{array}{lll}
    p \in \Lang(con(r^*)) &\iff& f(p) \in \Lang(r^*) \\
    \ryan{TODO} & & \\
  \end{array} \]

\end{proof}

\begin{defn}
If $r$ complies to state machine $M^A$, then we compile the policy $con(r)$ in the following way:

  \[ \begin{array}{ll}
    1. & \sigma^A(q,f(l)) = q' \iff \sigma(q,l) = q' \\
    2. & q \in F^A \iff q \in F \\
  \end{array} \]

That is, we construct a new automata over the same states and final states, but with a different transition function $\sigma$. This ensures that there is a bisimulation between the two state machines.

\end{defn}

\begin{lem}
  $M^A$ matches $f(p)$ $\iff$ $M$ matches $p$.
\end{lem}

\begin{proof}
  Suppose $M^A$ matches $f(p) = A_1, \dots, A_k$, then there exists a trace through $M^A$ of the form:

  $$q_0, q_1, \dots, q_{k+1}$$

  where $\sigma^A(q_i, A_i) = q_{i+1}$ and $q_{k+1} \in F^A$. We must show that $M$ matches $p = a_1, \dots, a_k$.

  From the construction of the automaton, we know that $\sigma(q_i, a_i) = q_{i+1}$ and $q_{k+1} \in F \iff q_{k+1} \in F^A$. Therefore, the trace $q_0, q_1, \dots, q_{k+1}$ is accepted by $M$. 
  %
  The same proof is symmetric for the other direction.
\end{proof}

\begin{thm}
  $M$ matches $p$ $\iff$ $p \in \Lang(con(r))$
  \begin{proof}
    \[ \begin{array}{lllr}
      M ~\text{matches}~ p &\iff& M^A ~\text{matches}~ f(p) & \textit{previous lemma} \\
                         &\iff& f(p) \in \Lang(r) & \textit{automaton construction} \\
                         &\iff& p \in \Lang(con(r)) & \textit{first lemma} \\
    \end{array} \]

  \end{proof}
\end{thm}

\subsection{Product Graph}

\begin{defn}
  Because compilation guarantees that the automata states are the same under the concretization function, we can lift the graph homomorphism $f$ to a new homomorphism $f_{pg}$ over the product graph in the following way:
  \[ \begin{array}{rcl}
    f_{pg}( s ) & = & s^A  \\
    f_{pg}( (l,q_1,\ldots,q_n) ) & = & (f(l),q_1,\ldots,q_n) \\
  \end{array} \]
\end{defn}

\begin{lem}
  If, for product graph $PG = (G,s,P)$ with $G=(V,E)$, if we have the following:

  \[ \begin{array}{ll}
    1. & (a,b) ~\text{is an edge in the concrete topology} \\
    2. & \tilde{m} = a \\
    3. & f_{pg}(m) = M \\
    4. & (M,N) ~ \text{is an edge in the product graph}~ PG^A \\
    5. & \tilde{N} = f(b) \\
  \end{array} \]

  then there exists a node $n$ in the concrete product graph $PG$ where:

  \[ \begin{array}{ll}
    1. & \tilde{n} = b \\
    2. & f_{pg}(n) = N \\
    3. & (m,n) ~ \text{is an edge in the product graph}~ PG \\
    4. & P(n) = P(N) \\
  \end{array} \]

  \begin{proof}

    suppose that $m = (a, q_1, \dots, q_k)$. Then we know that $M = (f(a), q_1, \dots, q_k)$. Also suppose there is some $b$ such that $N = (f(b), s_1, \dots, s_k)$ since $f^{-1}(N) \neq \emptyset$. This means that, since $(a,b)$ is an edge in the concrete topology, so $(f(a),f(b))$ is an edge in the abstract topology. Also we know that $\tilde{N} = f(b)$.

    Let us take the node $n = (b, s_1, \dots, s_k)$.

    \begin{easylist}
      & Clearly $\tilde{n} = b$
      & Clearly $f_{pg}(n) = (f(b), s_1, \dots, s_k) = N$
      & For each regular expression $\sigma^A(q_i, f(b)) = s_i$. This means that for each regular expression, $\sigma(q_i, b) = s_i$. It follows that $(m,n)$ is an edge in the product graph by definition.
      & Since the final states are the same for each state $q_i$ of each product graph state, it follows that $P(n) = P^A(f(n)) = P^A(N)$.
    \end{easylist}

  \end{proof}
\end{lem}

\begin{thm}
  $m \leq m'$ in the concrete product graph iff $f_{pg}(m) \leq f_{pg}(m')$ in the abstract product graph under the lifted graph homomorphism $f_{pg}$.

  \begin{proof}
    $(\leftarrow)$

    Suppose we have a simulation relation from the subgraphs rooted at nodes $M$ and $M'$ with $M' \geq_{rank} M$ in the abstract product graph. Assume that $\tilde{M} = f(a)$ and $\tilde{N} = f(b)$. For each neighbor $N'$, if we can make a move from $M'$:

    $$M' \overset{\tilde{N'}}{\rightarrow} N'$$

    then we also make a similar move for $M$ for some $N$ where $N \approx N'$ and $N' \geq_{rank} N$:

    $$M \overset{\tilde{N}}{\rightarrow} N$$

    We must show that the same relation holds for $m$ and $m'$.

    Now assume $m'$ has a neighbor $n'$ where $\tilde{m} = a$ and $\tilde{n} = b$. Then clearly $(a,b)$ is a valid topology edge since it is a transition in the product graph. We now apply the previous lemma with the following facts:

    \[ \begin{array}{ll}
      1. & (a,b) ~\text{is a topology edge} \\
      2. & \tilde{m} = a \\
      3. & f_{pg}(m) = M \\
      4. & (M,N) ~\text{is an edge in}~ PG^A \\
      5. & \tilde{N} = f(b) \\
    \end{array} \]

    This lets us conclude that there exists a node $n'$ such that:

    \[ \begin{array}{ll}
      1. & \tilde{n} = b \\
      2. & f_{pg}(n) = N \\
      3. & (m,n) ~\text{is an edge in}~ PG \\
      4. & P(n) = P(N) \\
    \end{array} \]

    Thus if we have $m' \overset{b}{\rightarrow} n'$, then we also have $m \overset{b}{\rightarrow} n$. The previous lemma can also be used to show that $P(n') = P(N')$ in a symmetric fashion, so we are guaranteed that $n' \geq_{rank} n$ from the fact that $N' \geq_{rank} N$.

  \end{proof}

  \begin{proof}
    $(\rightarrow)$

    This is the simpler case. We assume we have $m' \overset{b}{\rightarrow} n'$ implies $m \overset{b}{\rightarrow} n$.
    Simply take $N = f_{pg}(n)$ and $N' = f_{pg}(n')$. If we know that $M' \overset{b}{\rightarrow} N'$ then we already know that $M \overset{b}{\rightarrow} N$ from the homomorphism. All that's required is to show that $N' \geq_{rank} N$. This follows trivially from the fact that $P(n) = P(N)$ and $P(n') = P(N')$.
  \end{proof}

\end{thm}
